Parsing Postfix log files

### 2008-10-28 doc/logparser-acronyms.tex
First pass through corrections for sections 1-3.

\newacronym{SPF}{Sender Policy Framework}{description={
    SPF is introduced in \sectionref{spf introduction} and explained fully
    in~\cite{openspf, Wikipedia-spf}.
}}

@@

\newacronym{FQDN}{Fully Qualified Domain Name}{description={
    An FQDN is a hostname plus domain name, e.g.\ \newline
        example.com     is a domain name          \newline
        www             is a hostname             \newline
        www.example.com is a FQDN
}}

@@

\newacronym{regex}{Regular Expression}{description={
    Regular expressions are a concise and powerful method of matching
    patterns against text that are explained in detail
    in~\cite{Wikipedia-regex, perlre}.
}}

@@

\newacronym{HTML}{Hypertext Markup Language}{description={
    HTML is the markup language designed for writing web
    pages~\cite{Wikipedia-html, w3schools-html}.
}}

>>>>

\newacronym{SPF}{Sender Policy Framework}{description={
    SPF is introduced in \sectionref{spf introduction} and explained fully
    in~\cite{openspf,Wikipedia-spf}.
}}

@@

\newacronym{FQDN}{Fully Qualified Domain Name}{description={
    An FQDN is a hostname plus domain name, e.g.\ \newline{}
        example.com     is a domain name          \newline{}
        www             is a hostname             \newline{}
        www.example.com is a FQDN
}}

@@

\newacronym{regex}{Regular Expression}{description={
    Regular expressions are a concise and powerful method of matching
    patterns against text that are explained in detail
    in~\cite{Wikipedia-regex,perlre}.
}}

@@

\newacronym{HTML}{Hypertext Markup Language}{description={
    HTML is the markup language designed for writing web
    pages~\cite{Wikipedia-html,w3schools-html}.
}}

### 2008-10-28 doc/logparser-background.tex
First pass through corrections for sections 1-3.

This section provides background information helpful in understanding the
remainder of the document.  It begins with a discussion of the motivation
underlying the project, followed by some technical information: the use of
a database as an \API{}\@; a brief introduction to \SMTP{}\@; a longer
introduction to Postfix, concentrating on the topics most relevant to this
document, namely Postfix anti-spam restrictions and policy servers.  The
assumptions made in designing and implementing the parser are explained, as
are the conventions used in this document.

@@

This document and the program it describes are part of a larger project to
optimise a mail server's Postfix restrictions, generate statistics and
graphs, and provide a platform on which new restrictions can be trialled
and evaluated to determine if they are beneficial in the fight against
spam.  The program parses Postfix log files and populates a database with
the data gleaned from those log files, providing a consistent and simple
view of the log files that future tools can utilise.  The gathered data
can then be used to optimise current anti-spam measures, provide a baseline
to test new anti-spam measures against, or to produce statistics showing
how effective those measures are.

A snippet of \SQL{} provides a short example of the optimisation possible
using data from the database: determining which Postfix restrictions reject
the highest number of mails:

@@

Another example is determining which restrictions are not effective: this
shows which restrictions had fewer than 100 hits on the last log file
parsed, and the percentage of total rejections each restriction represents.

@@

These database queries yield summary statistics about the efficiency of
spam avoidance techniques; statistics that are far less feasible to assess
directly from log files without prior pre-processing into a database in the
fashion proposed, implemented and tested herein.

@@

The database populated by this program provides a simple interface to
Postfix log files.  Although the interface is a database schema, it is in
effect quite similar to any other \API{} provided by shared code: it
insulates both user and provider of the \API{} from changes in the
implementation of the \API{}\@.  The algorithm implemented by the parser
can be improved; support can be added for earlier or later releases of
Postfix; bugs can be fixed or limitations removed from the parser; these
changes will not cause the user to be negatively affected.  Statistics
and graphs can be generated from the database; new restrictions can be
tested and the results inspected; trends in the fight against spam can
emerge from historical data saved in the database; the parser remains the
same as the usage adapts.  Using a database simplifies writing programs
that need to interact with the data in several ways:

@@

    \item It is already possible to access databases from most programming
        languages, allowing a developer to access the data gathered using
        their preferred programming language, rather than being restricted
        to the language the parser is written in.  It is often possible to
        write an interface layer allowing code written in one language to
        be used in another language, but this greatly increases the effort
        required to use the parser.

@@

    \item \SQL{} is reasonably standard and many people will already be
        familiar with it; for those unfamiliar with it there are lots of
        readily available resources from which to learn (a good
        introduction to \SQL{} can be found at~\cite{sql-for-web-nerds},
        others are~\cite{w3schools-sql-tutorial, sqlcourse.com}).  Although
        every vendor implements a different dialect of \SQL{}, the basics
        are the same everywhere (analogous to the overall similarities and
        minor differences between Irish English, British English, American
        English and Australian English).  Depending on the database in use
        there may be tools available that reduce or remove the requirement
        to know \SQL{}; for SQLite (the default database used by the
        implementation) several are available~\cite{sqlite-guis}.

@@

Storing the results in a database will also increase the efficiency of
using those results, as the log files need only be parsed once rather than
each time the data is used; indeed the results may be used by someone with
no access to the original log files.

@@

The \SMTPlong{}, originally defined in \RFC{}~821~\cite{RFC821} and updated
in \RFC{}~2821~\cite{RFC2821}, is used for transferring mail between the
sending and receiving \MTA{}\@.  It is a simple, human readable, plain text
protocol, making it quite simple to test and debug problems with it.  Despite
the simplicity of the protocol many virus and spam sending programs fail
to implement it properly, so requiring strict adherence to the protocol
specification is beneficial in protecting against spam and
viruses.\footnote{\label{footnote:rfc760}Originally all mail servers
adhered to the principle of \textit{Be liberal in what you accept, and
conservative in what you send\/} from \RFC{}~760~\cite{rfc760}, but
unfortunately that principle was written in a friendlier time.  Given the
deluge of spam that mail servers are subjected to daily, a more appropriate
maxim could be: \textit{Require strict adherence to \RFC{}~2821; implement
the strongest restrictions you can; relax the restrictions and adherence
only when legitimate mail is impeded.\/}  it is not as friendly, nor as
catchy, but it more accurately reflects the current situation.} A typical
\SMTP{} conversation resembles the following (the lines starting with a
three digit number are sent by the server, all other lines are sent by the
client):

@@

This client is so poorly written that not only does it present the sender
and recipient addresses improperly, it ignores the error messages returned
by the server and carries on regardless.  There are many spam and virus
sending programs that are this deficient --- unfortunately others
(particularly newer programs) were written by competent programmers, or
utilise competently written programs (e.g.\ Postfix or Sendmail on Unix
hosts, Microsoft Outlook on Windows hosts).  Traditionally a mail server
would have done its best to deal with deficient clients, with the intention
of accepting as much mail destined for its users as
possible,\footref{footnote:rfc760} e.g.\ by ignoring the absence of a HELO
command, or accepting sender or recipient addresses that were not enclosed
in \texttt{<>}.

A detailed description of \SMTP{} is beyond the scope of this document:
introductory guides can be found at~\cite{smtp-intro-01, smtp-intro-02},
the definitive references are~\cite{RFC821, RFC2821}.

@@

Postfix is a \MTA{} with the following design aims (in order of
importance): security, flexibility of configuration, scalability, and high
performance.  It features extensive, extensible, optional anti-spam
restrictions, allowing an administrator to deploy those restrictions which
they judge suitable for their site's needs, rather than a fixed set chosen
by Postfix's author.  These restrictions can be selectively applied,
combined and bypassed on a per-client, per-recipient or per-sender basis,
allowing varying levels of stricture and permissiveness.  Postfix leverages
simple lookup tables to support arbitrarily complicated user-defined
sequences of restrictions and exceptions, with policy
servers\footnote{Policy servers will be explained in \sectionref{policy
servers}.} as the ultimate in flexibility.  Administrators can also supply
their own rejection messages to make it clear to senders why exactly their
mail was rejected.  Unfortunately this flexibility has a cost: complexity
in the log files generated.  Although it is easy to use standard Unix text
processing utilities to determine the fate of an individual email,
following the journey an email takes through Postfix can be quite
difficult.  For most mails the journey is simple and brief, but the
remaining minority can be quite complex (see \sectionref{additional
complications} for details).

Postfix's design follows the Unix philosophy of \textit{Write programs that
do one thing and do it well\/}~\cite{unix-philosophy}, and is separated
into various component programs to perform the tasks required of an
\MTA{}\@: receive mail, send mail, local delivery of mail, etc.\ --- full
details can be found in~\cite{postfix-overview}.  Each log line contains
the name of the Postfix component that produced it, and this information
is used when determining which rules should be used to parse each log line
(see \sectionref{rule characteristics} for details).  This design has
positive security implications also: those components that interact with
other hosts are not privileged,\footnote{Privilege in this case means the
power to perform actions that are limited to the administrator, and not
available to ordinary users.} so bugs in those components will not give an
attacker extra privileges; those components that are privileged do not
interact with other hosts, making it much more difficult for an attacker to
exploit any bugs that may exist in those components.

@@

Postfix uses one restriction list (containing zero or more restrictions)
for each stage of the \SMTP{} conversation: client connection, HELO
command, MAIL FROM command, RCPT TO commands, DATA command, and end of
data.  The appropriate restriction list is evaluated for each stage
(evaluation will be explained shortly), though by default the restriction
lists for client connection, HELO and MAIL FROM commands will not be
evaluated until the first RCPT TO command is received, because some clients
do not deal properly with rejections before RCPT TO\@; a benefit of this
delay is that Postfix has more information available when logging
rejections.

Each restriction is evaluated to produce a result of \textit{reject},
\textit{permit}, \textit{dunno\/} or the name of another restriction to be
evaluated.\footnote{Other results are possible as described
in~\cite{smtpd_access_readme, smtpd_per_user_control, policy-servers}.} The
meaning of \textit{permit\/} and \textit{reject\/} is fairly obvious;
\textit{dunno\/} means to stop evaluating the current restriction and
continue processing the remainder of the restriction list, allowing more
specific cases to be used as exceptions to more general cases.  When the
result is the name of another restriction Postfix will evaluate the new
restriction, allowing restrictions to be chosen based on the client \IP{}
address, client hostname, HELO hostname, sender address, or recipient
address.\footnote{E.g.\ the administrator may require that all clients on
the local network have valid \DNS{} entries, to prevent people sending mail
from unknown machines.}  The administrator can define new restrictions as a
list of existing restrictions, allowing arbitrarily long and complex
sequences of lookups, restrictions and exceptions.  Postfix tries to
protect the administrator in as far as is reasonable, e.g.\ the restriction
\texttt{check\_helo\_mx\_access} cannot cause a mail to be accepted,
because the parameter it checks (the hostname given in the HELO command) is
under the control of the remote client.  Despite this, it is possible for
the administrator to make catastrophic mistakes, e.g.\ rejecting mail for
all users --- the administrator must be cognisant of the effects of their
configuration changes.  This is similar to one of UNIX's design
philosophies: \textit{``UNIX was not designed to stop its users from doing
stupid things, as that would also stop them from doing clever things''\/}
--- Doug Gwyn~\cite{unix-philosophy}.

Postfix uses simple lookup tables to make decisions when evaluating some
restrictions, e.g.\ in the example restriction below\newline
\tab{}\texttt{check\_client\_access~cidr:/etc/postfix/client\_access}
\newline The line above can be broken down as follows:

\begin{description}

@@

\end{description}

The restriction \texttt{check\_client\_access} checks if the \IP{} address
of the connected client is found in the file and returns the associated
action if found; the method of searching the file is dependant on the type
of the file (\texttt{cidr} in the example) --- see
\cite{postfix-lookup-tables} for more details.  Other restrictions
determine their result by consulting external sources, e.g.\
\texttt{reject\_rbl\_client dnsbl.example.com} checks whether the \IP{}
address of the client is present in the \DNSBL{}
\texttt{dnsbl.example.com}, rejecting the command if the client is listed.
This description is necessarily brief, for further details
see~\cite{smtpd_access_readme, smtpd_per_user_control, policy-servers}.

@@

A \textit{policy server\/}~\cite{policy-servers} is an external program
consulted by Postfix to determine the fate of an \SMTP{} command.  The
policy server is given state information\footnote{Sample state information
is shown in \tableref{Example attributes sent to policy servers}} and
returns a verdict as described in \sectionref{Mixing and matching Postfix
restrictions}.  The policy server can perform more complex checks than
those provided by Postfix: a trivial example is restricting mail from
addresses associated with the payroll system to sending mail on the third
Tuesday after pay day only, to help prevent problems from spam or (worse)
phishing mails with faked sender addresses.\footnote{A phishing mail might
claim that the payroll system had a disastrous disk failure; until the
server is replaced and restored all salary payments will have to be
processed manually, so please reply to this mail with your name, address
and back account details.}

@@

    \item Checking \SPF{} records~\cite{openspf, Wikipedia-spf}.
        \SPF{}\label{spf introduction} records specify which mail servers
        are allowed to send mail claiming to be from a particular domain.
        The intention is to reduce spam from faked sender addresses,
        backscatter~\cite{postfix-backscatter} and
        joe~jobs~\cite{Wikipedia-joe-job}, but there has been considerable
        resistance to the proposal because it breaks or vastly complicates
        some features of \SMTP{}, e.g.\ forwarding mail from one company or
        university to another when a user changes jobs.

    \item Greylisting~\cite{greylisting} is a technique that temporarily
        rejects mail when the triple of (sender address, recipient address,
        remote \IP{} address) is unknown; on second and subsequent delivery
        attempts from that triple the mail will be accepted.  The
        assumption is that maintaining a list of failed addresses and
        retrying after a temporary failure is uneconomical for a spam
        sender, but that a legitimate mail server must retry.  Sadly spam
        senders are using increasingly complex and well written programs to
        distribute spam, frequently using an \ISP{} provided \SMTP{} server
        from a compromised machine on the \ISP{}'s network.  Greylisting
        will slowly become less useful, but it does block a large
        percentage of spam mail at the moment; the most effective
        restrictions over the \numberOFlogFILES{} log files used in testing
        the parser are shown in \tableref{Summary of rejections}.
        Greylisting is obviously worth using, at least at the moment,
        particularly when you consider Greylisting's position as the final
        restriction that a mail must overcome:\footnote{Greylisting is the
        final restriction a mail must overcome in the configuration used on
        the mail server the log files were obtained from; an administrator
        is free to use Greylisting at whichever position in the restriction
        list they feel is most appropriate for their mail system.}
        Greylisting only takes effect for mails that have passed every
        other restriction.

@@

    \item Using a scoring system such as
        Policyd-weight~\cite{policyd-weight} where tests accumulate points
        against the sending system --- if the eventual score is higher than
        a threshold the mail is rejected.  Postfix's restrictions are
        binary, and the administrator must manually whitelist clients if
        they are to bypass a restriction; using a threshold that requires a
        delivery attempt to hit several restrictions frees the
        administrator from whitelisting clients that fall foul of one
        restriction only.

@@

Example attributes taken from~\cite{policy-servers}:

    \begin{tabular}[]{ll}

    \end{tabular}
\end{table}

@@

@@

This section has provided background information on several topics relevant
to the remainder of the document.  It started with the motivation behind
the project, continuing with explanations of:

    \item Using a database as an \API{}.

    \item \SMTP{}.

    \item Postfix restrictions and policy servers.

    \item Assumptions made when designing and developing the parser.

    \item A description of the parser's novel design.

    \item Conventions used in this document.

    \item An in-depth comparison of this parser with \LMA{}, the parser
        described in previously published research.

    \item A review of the literature previously published in this area.

>>>>

This section provides background information helpful in understanding the
remainder of the document.  It begins with a discussion of the motivation
underlying the project, followed by some technical information: the use of
a database as an \API{}\@; a brief introduction to \SMTP{}\@; a longer
introduction to Postfix, concentrating on anti-spam restrictions and policy
servers.  The assumptions made in designing and implementing the parser are
explained, as are the conventions used in this document.

@@

This document and the program it describes are part of a larger project to
optimise a mail server's Postfix-based anti-spam restrictions, generate
statistics and graphs, and provide a platform on which new restrictions on
trial can be evaluated to determine if they are beneficial in the fight
against spam.  The program parses Postfix log files and populates a
database with the data gleaned from those log files, providing a consistent
and simple view of the log files that future tools can utilise.  The
gathered data can be used to optimise current anti-spam measures, to
provide a baseline to test new anti-spam measures against, or to produce
statistics showing how effective those measures are.

Determining which Postfix restrictions reject the highest number of mails
is a short example of the optimisation possible using data from the
database:

@@

Another example is determining which restrictions are not effective: this
sample query shows which restrictions had fewer than 100 rejections in the
last log file parsed, and the percentage of total rejections each of those
restrictions represents.

@@

The database queries above yield summary statistics about the efficiency of
anti-spam techniques.  These statistics are far less feasible to assess
directly from log files without prior pre-processing into a database in the
fashion proposed, implemented, and tested herein.

@@

The database populated by this program provides a simple interface to
Postfix log files.  Although the interface is a database schema, it is in
effect quite similar to any other \APIlong{}~(\API{}) provided by shared
code: it insulates both user and provider of the \API{} from changes in the
implementation of the \API{}\@.  The algorithm implemented by the parser
can be improved; support can be added for earlier or later releases of
Postfix; bugs can be fixed or limitations removed from the parser; these
changes will not cause the user to be negatively affected.  Statistics and
graphs can be generated from the database; new restrictions can be tested
and the results inspected; trends in the fight against spam can emerge from
historical data saved in the database; the parser remains the same as the
usage adapts.  Using a database simplifies writing programs that need to
interact with the data in several ways:

@@

    \item Facilities are provided to access databases from most programming
        languages, allowing a developer to access the data gathered using
        their preferred programming language, rather than being restricted
        to the language the parser is written in.  It is often possible to
        write an interface layer allowing code written in one language to
        be used in another language, but this greatly increases the effort
        required to use the parser.

@@

    \item \SQL{} is reasonably standard and many people will already be
        familiar with it; for those unfamiliar with it there are lots of
        readily available resources from which to learn (a good
        introduction to \SQL{} can be found at~\cite{sql-for-web-nerds},
        others are~\cite{w3schools-sql-tutorial,sqlcourse.com}).  Although
        every vendor implements a different dialect of \SQL{}, the basics
        are the same everywhere (analogous to the overall similarities and
        minor differences between Irish English, British English, American
        English, and Australian English).  Depending on the database in use
        there may be tools available that reduce or remove the requirement
        to know \SQL{}; several are available~\cite{sqlite-guis} for
        SQLite (the default database used by the parser
        implementation).

@@

Storing the results in a database will also increase the efficiency of
using those results, because the log files need only be parsed once rather
than each time the data is used; indeed the results may be used by someone
with no access to the original log files.

@@

The \SMTPlong{}, originally defined in \RFC{}~821~\cite{RFC821} and updated
in \RFC{}~2821~\cite{RFC2821}, is used for transferring mail between the
sending and receiving \MTA{}\@.  It is a simple, human readable, plain text
protocol, making it quite easy to test and debug problems with it.  Despite
the simplicity of the protocol many virus and spam sending programs fail
to implement it properly, so requiring strict adherence to the protocol
specification is beneficial in protecting against spam and
viruses.\footnote{\label{footnote:rfc760}Originally all mail servers
adhered to the principle of \textit{Be liberal in what you accept, and
conservative in what you send\/} from \RFC{}~760~\cite{rfc760}, but
unfortunately that principle was written in a friendlier time.  Given the
deluge of spam that mail servers are subjected to daily, a more appropriate
maxim could be: \textit{Require strict adherence to \RFC{}~2821; implement
the strongest restrictions you can; relax the restrictions and adherence
only when legitimate mail is impeded.\/}  it is not as friendly, nor as
catchy, but it more accurately reflects the current situation.} A typical
\SMTP{} conversation resembles the following (the lines starting with a
three digit number are sent by the server, all other lines are sent by the
client):

@@

This example client is so poorly written that not only does it present the
sender and recipient addresses improperly, it ignores the error messages
returned by the server and carries on regardless.  Many spam and virus
sending programs are this deficient --- unfortunately others (particularly
newer programs) were written by competent programmers, or utilise
competently written programs (e.g.\ Postfix or Sendmail on Unix hosts,
Microsoft Outlook on Windows hosts).  Traditionally a mail server would
have done its best to deal with deficient clients, with the intention of
accepting as much mail destined for its users as
possible,\footref{footnote:rfc760} e.g.\ by ignoring the absence of a HELO
command, or by accepting sender or recipient addresses that were not
enclosed in \texttt{<>}.

A detailed description of \SMTP{} is beyond the scope of this document:
introductory guides can be found at~\cite{smtp-intro-01,smtp-intro-02},
and the definitive references are~\cite{RFC821,RFC2821}.

@@

Postfix is a \MTAlong{}~(\MTA{}) with the following design aims (in order
of importance): security, flexibility of configuration, scalability, and
high performance.  It features extensive, extensible, optional anti-spam
restrictions, allowing an administrator to deploy those restrictions which
they judge suitable for their site's needs, rather than a fixed set chosen
by Postfix's author.  These restrictions can be selectively applied,
combined, and bypassed on a per-client, per-recipient, or per-sender basis,
allowing varying levels of stricture and permissiveness.  Postfix leverages
simple lookup tables to support arbitrarily complicated user-defined
sequences of restrictions and exceptions, with policy
servers\footnote{Policy servers will be explained in \sectionref{policy
servers}.} as the ultimate in flexibility.  Administrators can also supply
their own rejection messages to make it clear to senders why exactly their
mail was rejected.  Unfortunately this flexibility has a cost: complexity
in the log files generated.  Although it is easy to use standard Unix text
processing utilities to determine the fate of an individual email,
following the journey an email takes through Postfix can be quite
difficult.  For most mails the journey is simple and brief, but the
remaining minority can have quite complex journeys (see
\sectionref{additional complications} for details).

Postfix's design follows the Unix philosophy of \textit{``Write programs
that do one thing and do it well''\/}~\cite{unix-philosophy}, and is
separated into various component programs to perform the tasks required of
an \MTA{}\@: receive mail, send mail, local delivery of mail, etc.\ ---
full details can be found in~\cite{postfix-overview}.  Each log line
contains the name of the Postfix component that produced it, and this
information is used when determining which rules should be used to parse
each log line (see \sectionref{rule characteristics} for details).
Postfix's design is strongly influenced by security concerns: those
components that interact with other hosts are not
privileged,\footnote{Privilege means the power to perform actions that are
limited to the administrator, and not available to ordinary users.} so bugs
in those components will not give an attacker extra privileges; those
components that are privileged do not interact with other hosts, making it
much more difficult for an attacker to exploit any bugs that may exist in
those components.

@@

Postfix uses one restriction list (containing zero or more restrictions)
for each stage of the \SMTP{} conversation: client connection, HELO
command, MAIL FROM command, RCPT TO commands, DATA command, and end of
data.  The appropriate restriction list is evaluated for each stage
(evaluation will be explained shortly), though by default the restriction
lists for client connection, HELO and MAIL FROM commands will not be
evaluated until the first RCPT TO command is received, because some clients
do not deal properly with rejections before the first RCPT TO command; a
benefit of this delay is that Postfix has more information available when
logging rejections.

Each restriction is evaluated to produce a result of \textit{reject},
\textit{permit}, \textit{dunno\/} or the name of another restriction to be
evaluated.\footnote{Other results are possible as described
in~\cite{smtpd_access_readme,smtpd_per_user_control,policy-servers}.} The
meaning of \textit{permit\/} and \textit{reject\/} is fairly obvious;
\textit{dunno\/} means to stop evaluating the current restriction and
continue processing the remainder of the restriction list, allowing
exceptions to more general rules.  When the result is the name of another
restriction Postfix will evaluate that restriction, allowing restrictions
to be chosen based on the client \IP{} address, client hostname, HELO
hostname, sender address, or recipient address.\footnote{E.g.\ the
administrator may require that all clients on the local network have valid
\DNS{} entries, to prevent people sending mail from unknown machines.}  The
administrator can define new restrictions as a list of existing
restrictions, allowing arbitrarily long and complex sequences of lookups,
restrictions and exceptions.  Postfix tries to protect the administrator in
as far as is reasonable, e.g.\ the restriction
\texttt{check\_helo\_mx\_access} cannot cause a mail to be accepted,
because the parameter it checks (the hostname given in the HELO command) is
under the control of the remote client.  Despite this, it is possible for
the administrator to make catastrophic mistakes, e.g.\ rejecting all mail
--- the administrator must be cognisant of the effects of their
configuration changes.  This is similar to one of UNIX's design
philosophies: \textit{``UNIX was not designed to stop its users from doing
stupid things, as that would also stop them from doing clever
things''\/}~\cite{unix-philosophy}.

Postfix uses simple lookup tables to make decisions when evaluating some
restrictions, e.g.\newline{}
\tab{}\texttt{check\_client\_access~cidr:/etc/postfix/client\_access}

\begin{eqlist}

@@

\end{eqlist}

The restriction \texttt{check\_client\_access} checks if the \IP{} address
of the connected client is found in the specified file and returns the
associated action if found; the method of searching the file is dependant
on the type of the file (\texttt{cidr} in the example) --- see
\cite{postfix-lookup-tables} for more details.  Other restrictions
determine their result by consulting external sources, e.g.\newline{}
\tab{}\texttt{reject\_rbl\_client dnsbl.example.com}\newline{} checks the
\DNSBL{} \texttt{dnsbl.example.com}, rejecting the command if the client's
\IP{} address is listed.  XXX DO I NEED THIS SENTENCE\@?  This description
is necessarily brief, for further details see~\cite{smtpd_access_readme,
smtpd_per_user_control, policy-servers}.

@@

A policy server~\cite{policy-servers} is an external program consulted by
Postfix to determine the fate of an \SMTP{} command.  The policy server is
given state information by Postfix (sample state information is shown in
\tableref{Example attributes sent to policy servers}) and returns a verdict
from the set described in \sectionref{Mixing and matching Postfix
restrictions}.  The policy server can perform more complex checks than
those provided by Postfix: a trivial example is allowing addresses
associated with the payroll system to send mail on the third Tuesday after
pay day only, to help prevent problems from phishing mails using faked
sender addresses.\footnote{A phishing mail might claim that the payroll
system had a disastrous disk failure, and until the server is replaced all
salary payments will have to be processed manually, so please reply to this
mail with your name, address, and bank account details.}

    XXX HOW CAN I CENTRE THE TABLE\@?
    \begin{tabular}[]{ll}

    \end{tabular}

\end{table}

@@

    \item Checking \SPF{} records~\cite{openspf,Wikipedia-spf}.
        \SPF{}\label{spf introduction} records specify which mail servers
        are allowed to send mail claiming to be from a particular domain.
        The intention is to reduce spam from faked sender addresses,
        backscatter~\cite{postfix-backscatter}, and
        joe~jobs~\cite{Wikipedia-joe-job}.  There has been considerable
        resistance to the proposal because it breaks or vastly complicates
        some features of \SMTP{}, e.g.\ forwarding mail from one company or
        university to another when a user moves.

    \item Greylisting~\cite{greylisting} is a technique that temporarily
        rejects a delivery attempt when the triple of \newline{}
        \tab{}\texttt{(sender address, recipient address, remote \IP{}
        address)}\newline{} has not been seen before; on second and
        subsequent delivery attempts from that triple the mail will be
        accepted.  This blocks spam because maintaining a list of failed
        addresses and retrying after a temporary failure is uneconomical
        for a spam sender, but a legitimate mail server must retry
        deliveries that temporarily failed.  Sadly spam senders are using
        increasingly complex and well written programs to distribute spam,
        frequently using an \ISP{} provided \SMTP{} server from a
        compromised machine on the \ISP{}'s network.  Greylisting will
        slowly become less useful, but it does block a large percentage of
        spam mail at the moment; the most effective restrictions from the
        \numberOFlogFILES{} log files used in testing the parser are shown
        in \tableref{Summary of rejections}.  Greylisting is obviously
        worth using, at least at the moment, particularly when you consider
        Greylisting's position as the final restriction that a mail must
        overcome:\footnote{Greylisting is the final restriction a mail must
        overcome in the configuration used on the mail server the log files
        were obtained from; an administrator is free to use Greylisting
        wherever in the restriction list they feel is most appropriate for
        their mail system.} Greylisting only takes effect for mails that
        have passed every other restriction.  Some problems may be
        encountered when using greylisting: some servers fail to retry;
        legitimate mail may be delayed, particularly when coming from a
        pool of servers.

@@

    \item Scoring systems such as Policyd-weight~\cite{policyd-weight}
        perform tests on features of the delivery attempt (e.g.\ \IP{}
        address, sender address), incrementing or decrementing a score
        based on the results; if the eventual score is higher than a
        threshold the mail is rejected.  The administrator must manually
        whitelist clients if they are to bypass a Postfix restriction;
        using a threshold that requires a delivery attempt to hit several
        restrictions frees the administrator from whitelisting clients that
        fall foul of one restriction only.

@@

@@

XXX MAYBE REWRITE SO THAT THIS IS NOT TRUE\@?  ARE THERE ANY OTHER
CONVENTIONS\@?

@@

This section has provided background information on several topics,
starting with the motivation behind the project, continuing with an
explanation of the use of a database as an \API{} and an introduction to
\SMTP{}.  The focus then switched to the \MTA{} in use, Postfix, and its
restrictions and policy servers, finishing with the conventions used in
this thesis.

### 2008-10-28 doc/logparser-implementation.tex
First pass through corrections for sections 1-3.

    \item [result\_cols, connection\_cols] Specifies how the fields in the
        log line will be extracted.  The format is: \newline \tab{}
        \texttt{smtp\_code = 1; recipient = 2, sender = 4;} \newline i.e.\
        semi-colon or comma separated assignment statements, with the
        variable name on the left and the matching capture from the
        \regex{} on the right hand side.  Data specified by
        \texttt{result\_cols} will be entered in the \texttt{results}
        table, whereas \texttt{connection\_cols} populates the
        \texttt{connection} table.  The list of acceptable variable names
        is:

        \texttt{connection\_cols: client\_hostname, client\_ip, server\_ip,
        \newline \tab{} server\_hostname} and \texttt{helo.\newline}
        \texttt{result\_cols: sender, recipient, smtp\_code, message\_id,
        \newline \tab{} size,} and \texttt{data}

@@

    \item [name, description, restriction\_name and postfix\_action:] are
        not \newline used by the algorithm, they serve to document the rule
        for the user's benefit.

@@

    \item [result\_cols, connection\_cols, result\_data and
        connection\_data:] are \newline used by the action to extract and
        save data matched by the \regex{}.

@@

        \texttt{3FF7C4317: to=<username@example.com>, relay=local, \newline
        delay=0, status=sent (forwarded as 56F5B43FD)}

@@

    \item The message-id has a specific format: \newline
        \texttt{YYYYMMDDhhmmss.queueid@server.hostname} \newline
        e.g.\ \texttt{20070321125732.D168138A1@smtp.example.com}

@@

    \item [\daemon{local}, \daemon{pickup}, \daemon{smtp}:] Locally
        submitted mail, \newline both local and remote delivery.

@@

            \item Create \texttt{/etc/postfix/log\_helo.pcre}
                containing:\newline \tab{}\texttt{/\^/~~~~WARN~Logging~HELO}

            \item Modify \texttt{smtpd\_data\_restrictions} in
                \texttt{/etc/postfix/main.cf} to contain\newline
                \tab{}\texttt{check\_helo\_access~/etc/postfix/log\_helo.pcre}

>>>>

    \item [result\_cols, connection\_cols] Specifies how the fields in the
        log line will be extracted.  The format is: \newline{} \tab{}
        \texttt{smtp\_code = 1; recipient = 2, sender = 4;} \newline{} i.e.\
        semi-colon or comma separated assignment statements, with the
        variable name on the left and the matching capture from the
        \regex{} on the right hand side.  Data specified by
        \texttt{result\_cols} will be entered in the \texttt{results}
        table, whereas \texttt{connection\_cols} populates the
        \texttt{connection} table.  The list of acceptable variable names
        is:

        \texttt{connection\_cols: client\_hostname, client\_ip, server\_ip,
        \newline{} \tab{} server\_hostname} and \texttt{helo}.\newline{}
        \texttt{result\_cols: sender, recipient, smtp\_code, message\_id,
        \newline{} \tab{} size,} and \texttt{data}

@@

    \item [name, description, restriction\_name and postfix\_action:] are
        not \newline{} used by the algorithm, they serve to document the rule
        for the user's benefit.

@@

    \item [result\_cols, connection\_cols, result\_data and
        connection\_data:] are \newline{} used by the action to extract and
        save data matched by the \regex{}.

@@

        \texttt{3FF7C4317: to=<username@example.com>, relay=local, \newline{}
        delay=0, status=sent (forwarded as 56F5B43FD)}

@@

    \item The message-id has a specific format: \newline{}
        \texttt{YYYYMMDDhhmmss.queueid@server.hostname} \newline{}
        e.g.\ \texttt{20070321125732.D168138A1@smtp.example.com}

@@

    \item [\daemon{local}, \daemon{pickup}, \daemon{smtp}:] Locally
        submitted mail, \newline{} both local and remote delivery.

@@

            \item Create \texttt{/etc/postfix/log\_helo.pcre}
                containing:\newline{} \tab{}\texttt{/\^/~~~~WARN~Logging~HELO}

            \item Modify \texttt{smtpd\_data\_restrictions} in
                \texttt{/etc/postfix/main.cf} to contain\newline{}
                \tab{}\texttt{check\_helo\_access~/etc/postfix/log\_helo.pcre}
